---
title: "BSMM-lab-2"
subtitle: "BSMM 8740 Fall 2023"
author: "Yamini Ponnambalam"
date: "25 Sep 2023"
format: html
editor: visual
self-contained: true
---

## Setup

Load packages and data:

```{r load-pkg-data}
#| message: false
the_tate <- readr::read_delim("data/the-tate-collection.csv", ";", escape_double = FALSE, trim_ws = TRUE)
the_tate_artists <- readr::read_csv("data/the-tate-artists.csv")
```

```{r}
library(magrittr)     # the pipe
library(tidyverse)    # for data wrangling + visualization
library(tidymodels)   # for modeling
library(gt)           # for making display tables
library(gtExtras)     # helper functions for beautiful tables
library(DataExplorer) #
```

## Exercises

### Exercise 1

```{r}
the_tate <- readr::read_delim("data/the-tate-collection.csv", ";", escape_double = FALSE, trim_ws = TRUE)
```

```{r}
the_tate_artists <- readr::read_csv("data/the-tate-artists.csv")
```

```{r}
 summarizing_tate_artists <- the_tate |>
  dplyr::summarize(
    unique_artists = n_distinct(artist),
    min_date = min(year, na.rm = TRUE), 
    max_date = max(year, na.rm = TRUE),
    min_acquisition_year = min(acquisitionYear, na.rm = TRUE),
    max_acquisition_year = max(acquisitionYear, na.rm = TRUE)
    )
 summarizing_tate_artists
```

```{r}
DataExplorer::introduce(the_tate)
```

```{r}
DataExplorer::plot_missing(the_tate)
```

The `the_tate` dataset has \_3336\_\_ unique artists who worked from \_1545\_\_ to \_2012\_\_. The works were acquired between the years \_1823\_\_ and \_2013\_\_.

### Exercise 2

```{r}
the_missing_date_tates <- the_tate |>
  filter(is.na(year)) 
the_missing_date_tates
```

```{r}
total_works_with_missing_dates <- nrow(the_missing_date_tates)
total_works_with_missing_dates

```

```{r}
missing_dates_by_artist <- table(the_missing_date_tates)
missing_dates_by_artist_tibble <- as_tibble(missing_dates_by_artist) |>
  rename(artist = .rownames, missing_dates_count = n) |>
  arrange(desc(missing_dates_count))

missing_dates_by_artist
```

How number of works with missing dates is \_5397\_.

The number of artists whose works have missing dates is \_\_.

It would require resolving missing year data for only \_\_ artists resolve resolve at least 50% of the missing data.

The missing year data likely to be classified as \_\_\_\_.

### Exercise 3

The artist with the most works in the Tate collection is \_\_\_.

The artist with the tenth-most works in the Tate collection is \_\_\_.

### Exercise 4

The artist with the greatest number of works in the Tate collection represent \_\_\_% of the total number of works

### Exercise 5

There are \_\_ duplicate artist-title pairs

### Exercise 6

The artist with the largest work in the tate collection is \_\_\_

The artist with the smallest work in the collection is \_\_\_. The smallest work has area \_\_\_ $\text{cm}^2$

### Exercise 7

...

### Exercise 8

The annual return in the SPX price in 2020 was \_\_\_%.

The corresponding price volatility was \_\_\_%.

### Exercise 9

The period volatility was \_\_\_.\_%

### 
